TADCompare: an R package for differential analysis and characterization of Topologically Associated Domains

### 2019-12-17 manuscript/manuscript.Rmd
Add BMC Bioinformatics submission

---
bibliography: [references/networks.bib,references/brain.bib,references/enrichment.bib,references/3D_refs.bib]
csl: styles.ref/bmc-bioinformatics.csl
output:
  word_document:
    reference_docx: styles.doc/BMC_Bioinformatics.docx
  pdf_document: default
  html_document: default
editor_options: 
  chunk_output_type: console
---

@@

We collected 133 phenotype-associated gene lists from the NHGRI-EBI GWAS catalog [@MacArthur:2017aa] (see Methods). They were grouped into seven disease categories (“Autoimmune”, “Cancer”, “Cardiovascular disease”, “Eye disease”, “Metabolic disease”, “Neurologic”, “Psychiatric”) and five trait categories (“Anthropometric trait”, “Cardiovascular trait”, “Eye trait”, “Intelligence”, “Metabolic trait”, Table 1, Supplementary Table 1).

@@

@@

@@

We estimated coherence of 49 traits from five categories (Figure 3, Supplementary Table S3, Supplementary Figures S2-S3). Among anthropometric traits, Weight and Body Mass Index had the highest level of coherence. Other examples of traits with high coherence include blood pressure-related traits (cardiovascular) and glucose-related traits (metabolic). Notably, the metabolic trait category had the highest overall coherence (mean coherence 0.86, Table 1), and these results were consistent when using other PPI databases (Supplementary Table S1). Several notable examples of low coherence include Obesity-related traits, Height (anthropometric traits), and blood count-related traits (cardiovascular). Traits in the Intelligence category had low overall coherence (mean coherence = 0.41).

@@

@@

We further estimated coherence of 67 diseases from seven categories (Figure 4, Supplementary Figures S4-S5, Supplementary Table S3). Metabolic diseases had high overall coherence (average coherence 0.91, Table 1), with "Obesity" and "Type 2 diabetes" being the most coherent (1.01 and 0.91, respectively). Similarly, the average coherence for cancer diseases was 0.72, with "Testicular germ cell tumor" and "Pancreatic cancer" being highly coherent (0.92 and 0.74, respectively).

@@

@@

>>>>

---
bibliography: [references/networks.bib,references/brain.bib,references/enrichment.bib,references/3D_refs.bib]
csl: styles.ref/bmc-bioinformatics.csl
output:
  word_document:
    reference_docx: styles.doc/Arial_11_single_space_normal_margins.docx
  pdf_document: default
  html_document: default
editor_options: 
  chunk_output_type: console
---

@@

We collected 133 phenotype-associated gene lists from the NHGRI-EBI GWAS catalog [@MacArthur:2017aa] (see Methods). They were grouped into seven disease categories (“Autoimmune”, “Cancer”, “Cardiovascular disease”, “Eye disease”, “Metabolic disease”, “Neurologic”, “Psychiatric”) and five trait categories (“Anthropometric trait”, “Cardiovascular trait”, “Eye trait”, “Intelligence”, “Metabolic trait”, Table 1, Supplementary Table S1).

@@

<!--
**Figure 1. Degree distributions are affected by network size and configuration.** Degree distributions of A) ring networks and B) fully connected networks containing 10 and 20 nodes. While degree distributions of similarly coherent ring networks can be directly compared, degree distributions of fully connected networks depend on network size.
-->

@@

<!--
**Figure 2. Network size is inversely associated with coherence.** Slopes of internal vs. external degree distributions (aka coherence) for KEGG and random networks are plotted against network size.
-->

@@

We estimated coherence of 49 traits from five categories (Figure 3, Supplementary Table S3, Supplementary Figure S2, 6). Among anthropometric traits, Weight and Body Mass Index had the highest level of coherence. Other examples of traits with high coherence include blood pressure-related traits (cardiovascular) and glucose-related traits (metabolic). Notably, the metabolic trait category had the highest overall coherence (mean coherence 0.86, Table 1), and these results were consistent when using other PPI databases (Supplementary Table S1). Several notable examples of low coherence include Obesity-related traits, Height (anthropometric traits), and blood count-related traits (cardiovascular). Traits in the Intelligence category had low overall coherence (mean coherence = 0.41).

@@

<!--
**Figure 3. Coherence estimates of traits.** Size of dots represents the level of normalized coherence (X-axis) for individual traits (Y-axis).
-->

@@

We further estimated coherence of 67 diseases from seven categories (Figure 4, Supplementary Figure S4, 8, Supplementary Table S3). Metabolic diseases had high overall coherence (average coherence 0.91, Table 1), with "Obesity" and "Type 2 diabetes" being the most coherent (1.01 and 0.91, respectively). Similarly, the average coherence for cancer diseases was 0.72, with "Testicular germ cell tumor" and "Pancreatic cancer" being highly coherent (0.92 and 0.74, respectively).

@@

<!--
**Figure 4. Coherence estimates of diseases.** Size of dots represents the level of normalized coherence (X-axis) for individual diseases (Y-axis).
-->

@@

<!--
# Supplementary material

**Supplementary Figure S5. Coherence estimates of diseases using STRING as a reference PPI database.** See legend for Supplementary Figure S2.
-->
